Astrophysical calibration of gravitational-wave detectors

### 2015-10-08 docs/cosmic_cal.tex
Finished my final edits before LVC circulation.

% introduce GWs and BNS systems
It is expected that the advanced generation of interferometric \ac{GW} detectors will detect waves 
emitted from $O(10s)$ of \acp{CBC}. One such class of these cataclysmic events, the inspiral and
merger of \ac{BNS} systems will be detected out to a maximum range of $\approx 450$ Mpc. Assuming 
the current best estimates for the cosmological parameters, this is equivalent to a redshift 
$z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC} systems can be used as cosmological 
distance markers, otherwise known as ``standard sirens'' (analogous to the \ac{EM} standard 
candles). \cm{Standard sirens allow ... fix me}

% BNS systems are GRBs
It is considered likely that the merger of \ac{BNS} and/or \ac{NSBH} systems, in addition to 
emitting detectable \acp{GW}, is also the mechanism for producing \acp{sGRB} 
\cite{1992ApJ...395L..83N}. In this scenario these events produce tightly beamed \ac{EM} emission 
parallel to the orbital angular 
momentum vector of the \ac{BNS} system. Additional evidence for the coincidence of \acp{sGRB} with
\ac{GW} events are the estimated astrophysical rates of both phenomena. Observations of \acp{sGRB} 
give a rate of \cm{XXX}\,Mpc$^{-3}$\,yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger 
rates of $10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or \ac{NSBH} merger 
rates of $6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$ 
\cite{2010CQGra..27q3001A}. However, the recent work of \cite{2015MNRAS.448..928K}, which studied 
known galactic \ac{BNS} systems in more detail suggest a rate approximately five times lower. The 
estimates for \ac{NSBH} systems come entirely from population synthesis models as no such systems 
have yet been observed.
%The ranges of rates for \acp{BNS} is obtained from population synthesis models together with 
%knowledge of the X known \ac{BNS} systems in our galaxy, whilst the range for \acp{NSBH} is 
%entirely from population synthesis as not such systems have been observed.

@@

\MP{Move this paragraph to conclusions} For future planned space-based detectors such as eLISA there are several known galactic binary 
systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that their masses and 
orbital parameters allow them to be guarenteed sources for such detectors. These too could be used 
for assessing calibiration for such detectors. However, generally for these verification binaries 
the inclination of the system is not known and distances are fairly uncertain, so an accurate 
calibration assessment may not be possible. Although there is at least one (J0651) currently known 
eclipsing white dwarf binary \cite{2041-8205-737-1-L23}, which does provide the system inclination 
\cite{2012ApJ...757L..21H, 2013ASPC..467...47K} and has an approximately 10\% distance uncertainty 
\cite{2041-8205-737-1-L23}, which make it an excellent candiate for an astrophysical calibrator at 
its frequency.

@@

% Standard siren basics
The idea of \ac{GW} standard sirens is directly analogous to the concept of
standard candles in \ac{EM} astronomy and was first proposed
in~\cite{1986Natur.323..310S}. Unlike the primary \ac{EM} standard candle
event, Type Ia Supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
not only a function of distance. It also depends upon the chirp mass (a
function of the component masses) and the binary orientation with respect to
the detector network. However, measurement of the phase evolution of such an
event allows accurate determination of the chirp mass~\footnote{In reality it
is the redshifted chirp mass that is measured.}. Timing information from the
different signal arrival times at each detector in the network allows
reasonably accurate sky position determination.  Finally, amplitude variation
between differently oriented interferometers allows some (weaker) level of
determination of the binary system orientation.
%A great help to this final
%stage for this project is the assumption that the source has an \ac{sGRB}
%counterpart.  It follows that we therefore have a well constrained prior belief
%on the inclination of the source orbit with respect to the line-of-sight based
%on the expected level of beaming in the \ac{sGRB} emission.

% The classic use of sirens
Standard sirens are clearly a very powerful tool for \ac{GW} cosmology since
they give us a direct measure of the absolute luminosity distance to sources.
This is distinct from Type Ia Supernovae standard candles that only provide
relative luminosity distance measures and require calibration via other methods
as part of the cosmological distance ladder. At first sight the missing
component for standard sirens is a complementary measurement of the source
redshift which would allow each event to inform the distance-redshift
relationship and hence obtain cosmological parameter estimates.  However, it
was not thought possible (until recently) that redshift measurements could be
obtained from \ac{GW} observations. In~\cite{DelPozzo:2011uf} (building on the
original idea proposed in~\cite{1986Natur.323..310S}) the idea of identifying
host galaxies within the sky position uncertainty region was developed. Each
galaxy would provide a redshift measure and would contribute to a disparate mix
of possible cosmological parameters. However, by combining multiple (${\sim}20$)
detections, the true set of cosmological parameters emerges with an expected
${\sim}5\%$ uncertainty on the Hubble constant.

% continued use of sirens
\cm{Lose, trim or merge this paragraph with previous one.} The use of \ac{GW} events with \ac{sGRB} counterparts for use as a cosmological tool was 
investigated in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
the 3rd generation \ac{GW} interferometer, the \ac{ET}. It was
assumed in both studies that 1000 coincident events would be observed and
cosmological parameter estimation ability was compared with current \ac{EM}
results with focus on dark energy parameters. Other methods for cosmological
inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T, 2012PhRvD..86b3502T}) have 
been proposed for \ac{BNS} systems that do
not rely on any redshift measurements and instead use the distribution of
measured \acp{SNR} combined with some assumptions regarding the \ac{NS} mass
and spatial distributions. Finally, we mention that for \ac{BNS} systems it has
been shown in~\cite{Messenger:2011ux} and~\cite{2013arXiv1312.1862M} that features
of the tidal and post-merger hyper-massive \ac{NS} stages of the \ac{GW}
waveform can allow the measurement of the source redshift if the \ac{NS}
equation-of-state is sufficiently well known. This development, and separately, the method
based on \ac{SNR} distributions will allow \ac{GW} cosmological inference
without the need for complementary \ac{EM} observations.

@@

% discuss joint-event rates
The predicted detection rates of \ac{BNS} events (irrespective of an \ac{sGRB}
counterpart) are derived from three main sources, population synthesis models,
extrapolations based on known galactic \ac{BNS} systems, and from observations
of \ac{sGRB} events. Various studies using the first two methods have been
compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{\cm{Email Patrick
S about this} We note that this rate is that given 
in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO} and \ac{AdV} network consisting to two 
\ac{aLIGO} detectors operating at design sensitivity. However, a higher rate of $0.4$--$400$ is 
given in \cite{2010CQGra..27q3001A}, which may differ due to the use of a third \ac{aLIGO} 
detector, but the discrepancy may also have contributions from slightly different threshold 
definitions or design sensitivity curves used. We also note the study mentioned in 
Sec.~\ref{sec:intro} that potentially lowers the expected rate by a factor of five.}. 
In~\cite{2012MNRAS.425.2668C} using the rate of observed \acp{sGRB} and
assuming a beaming angle of $15^{o}$ a similar event rate of $1$--$180$ is
obtained. The beaming angle is the primary factor in determining the fraction
of detected \ac{GW} \ac{BNS} signals with an \ac{EM} \ac{sGRB} counterpart and
is unfortunately highly uncertain.  Various authors have provided estimates on
the rate of joint detections to be in the range $0.02$--$7$ per
year~\cite{2013PhRvL.111r1101C,Kelley:2012fl,2014arXiv1405.2254S,2015MNRAS.448.3026W}.

@@

% the main aim of the analysis
The main aim of this work is to assess how well the calibration scale factor
defined in Eqn.~\ref{eq:scalefactor} can be estimated from a single observed
\ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have
little knowledge of the likely location and distance of such an event we have
performed simulations of multiple events to see the distribution in the accuracy of the calibration 
scale factor recovery.

@@

% mention of distance delta function prior
The primary reason why coincident observations with a \ac{sGRB} can be used as
a check of detector calibration is that the \ac{sGRB} allows us to constrain
the priors on various parameters that would normally be highly correlated with
the signal amplitude. The most important of these is that the \ac{sGRB} can
provide a very tight constraint on the source distance. In this analysis we therefore
assume that the source distance is known (i.e.\ has a $\delta$-function
prior).~\cm{Should we not say a little bit more about this? A really accurate
distance from a really accurate redshift is only as good as the Hubble constant
which currently has $\sim$ few percent error?  We should also say something
about peculiar motion.}

@@

% discuss the component mass and spin priors 
The prior on the system component masses is dependent on whether we are
considering the source being a \ac{BNS} system or a \ac{NSBH}. For the former
case the prior we use is a Gaussian distribution for both components with means
of $1.35\,\textrm{M}_{\odot}$ and standard deviations of
$0.13\,\textrm{M}_{\odot}$, which is roughly consistent with the double neutron star system
masses found in \cite{2012ApJ...757...55O} albeit with a wider distribution. In reality, for all 
the systems we use in our study, the chirp mass, which we see from Eq.~\ref{eq:signal}
plays a role in the overall signal amplitude, is very well constrained by the
system's phase evolution, so our prior could be expanded with minimal effect on
the results. In the case of \ac{NSBH} systems we use the same prior as above
for the neutron star, but for the black hole we use a prior based on the
canonical masss distribution used for the rate results in
\cite{2012PhRvD..85h2002A} with a mean of $5\,\textrm{M}_{\odot}$ and a
standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass range of
black holes could be quite different from this, but use this range as an
example for this type of system. For the \ac{NSBH} systems we also require a
prior on the black hole spin (for \ac{BNS} systems we have assumed that the
spins are small enough that they will be negligible). We use a uniform prior on
the normalised aligned spin magnitude between $-1$ and $1$.

@@

% Describe the ensemble of simulations used
Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
number (several hundred up to $O(1000)$) simulations at each distance value. When
proposing injections for each simulation the source parameters were randomly
drawn from the prior distributions described in Section~\ref{sec:priors}, with
the sky position being drawn randomly from a uniform distribution on the sky,
the coalescence times held fixed within the centre of the prior window, and
with the calibration scaling factors for each detector drawn from a Gaussian
distribution with a mean of one and standard deviation of $0.125$ (equivalent
to a mean fractional offset of 10\%)~\cm{should we explicitly acknowledge that
we do not simulate from the same priors that we assume in the recovery?}.
However, in accepting a proposed injection as one to be analysed we introduced
a criteria that the signal be ``detectable'', based on that used in
\cite{2012PhRvD..85h2002A}. We therefore only analysed injections that had an
\ac{SNR} of $\geq 5.5$ in at least two detectors. Since we assume that these
signals would be coincident with a \ac{sGRB} we do not include the often used
further constraint that the total coherent network \ac{SNR} is greater than 12.
For injections close to the \ac{SNR} threshold, the application of the
threshold resulted in the preferential selection of better oriented systems.
Therefore, at larger distances the population of injected sources was not
uniformly distributed over the sky and instead favoured locations with better
antenna response.  Similarly, despite being already constrained by the prior to small
inclination angles, the accepted injections show preference for sources very
close to circularly polarised.

@@

% describe the main boxplot results
From these posterior distributions we have calculated the minimal 68\% credible
region for the calibration scale factors for each detector (if these were
Gaussian distributions this is equivalent to the region either side of the mean
bounded by the $1\sigma$ intervals). For all the signals at each distance
increment we have produced the distribution of the fractional half widths
(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the
true value. These are shown as boxplots in figures~\ref{fig:bnsresults} and
\ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems respectively. The
boxes show the extent from the lower to upper quartile of the values, whilst
the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The
black line within each box gives the median value and the star gives the mean
value. Also shown on each plot as the dashed magenta line is the percentage of
signals drawn from the prior distribution that fulfil the detection criterion.

@@

% discuss the BNS plots - general behaviour
In figure~\ref{fig:bnsresults} we see that on average the scale factor can be
recovered to equivalent precision in both \ac{aLIGO} detectors, as would be
expected, with uncertainties generally within the 10\% range for sources at
100\,Mpc. This is comparable to previous estimates of the calibration error in
the initial LIGO detectors. An interesting feature is that for distances
$\gtrsim 250$\,Mpc the upper extent of the uncertainty for H1 and L1 hits a
maximum at $\sim 25\%$, whilst the width of boxes narrows. This is due to our
``detectability'' criteria, where for all distances we will only be seeing
those sources with a high enough \ac{SNR} that we would consider them
detectable. There will therefore be fewer sources with \ac{SNR} higher than
this criteria at large distances, giving us a narrower range, and we
automatically exclude those with \ac{SNR} that are too small thus truncating
our uncertainty distribution at the upper end.  However, this does show that on
average for sources that are detectable out to 450\,Mpc we would be able to
constrain the calibration scale factor uncertainty for the \ac{aLIGO} detectors
to $\sim 20\%$.

@@

% summary of main results
We have established how well we can assess the detector calibration for
\ac{aLIGO} and \ac{AdV} using astrophysical sources as standard sirens. To do
this we require that a \ac{sGRB}, for which it has been possible to measure the
distance, is observed in coincidence with a \ac{CBC} signal in the \ac{GW}
detectors. This enables us to assume a known distance and sky position for the
source, and also limit the inclination of the source, which in turn allows us
to test the consistency of the detector calibration. We do this by including an
unknown scale factor on the true signal and noise, which we estimate given the
data from the 3 inteferometer advanced detector network. We find that for
detectable \ac{BNS} sources the uncertainty on the calibration scale factor
could on average be determined to $\lesssim 10\%$ of its true value for the
\ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the
proposed accuracy level of the hardware calibration of the detectors. For
sources at the standard single detector \ac{BNS} horizon distance of $\sim
450$\,Mpc the scale factor could on average be determined to within $\lesssim
20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were
found for \ac{NSBH} sources, although sources could be observed out to higher
distances.

@@

% comparitive calibration
%We imagine that a test for consistency of calibration between detectors could be
%performed with an ensemble of \ac{CBC} sources without relying on \ac{EM}
%counterparts. Such a method would have to marginalise over highly
%uncertain population models and may give quite weak bounds on the overall
%calibration uncertainty.~\cm{I'm not sure what you mean here? I want to add a
%paragraph on relative calibration after I understand what you mean.}

\cm{Chris: feel free to add a short paragraph on relative calibration here.}

% discus PSD uncertainty
We emphasise that our work has assumed that the \ac{PSD} used in the likelihood
function is an accurate representation of the \ac{PSD} at the time of the
observed \ac{CBC} signal. In reality the \ac{PSD} may be estimated from a
period close to, but not overlapping, the signal time and therefore may be
slightly different \cite{2013PhRvD..88h4044L}. Our result would therefore be
highly correlated with any uncertainty in the \ac{PSD} estimate, but would
still offer an upper limit on the calibration uncertainty. 
%However, future
%plans for \ac{CBC} parameter estimation include estimating the \ac{PSD} with
%the signal using BayesLine \cite{2015PhRvD..91h4034L}~\cm{I'm nit sure that
%this last point really helps us.}.

% frequency dependence
A further extension of this work would be for the calibration scale factor to
be estimated as a function of frequency. This may require the frequency series
to be divided into sections and the calibration scale factor estimated for
each. However, since the scale factor uncertainty is related to the \ac{SNR}
then the \ac{SNR} contribution in each frequency section will be of relevance.
In this case Bayesian model selection may be applicabl for selecting the
optimal number of frequency sections with which to estimate the scale factor.
The BayesLine algorithm \cite{2015PhRvD..91h4034L}, or the method described in
\cite{2013PhRvD..88h4044L}, may provide a natural way to perform such an
analysis. Indeed these methods, and those of \cite{Vitale:2012} are already
being used for marginalising over calibration uncertainties, but do not yet
attempt to estimate it.

@@

We would like to acknowledge the useful discussions with a whole bunch of people. LW was part 
funded for this work through a Royal Astronomical Society Undergraduate Research Bursary.
MP is funded by the STFC under grant number ST/L000946/1 and CM is funded via
XXX. This document has LIGO DCC number \dcc.

>>>>

% introduce GWs and BNS systems
It is expected that the advanced generation of interferometric \ac{GW}
detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year. One
such class of these cataclysmic events, the inspiral and merger of \ac{BNS}
systems will be detected out to a maximum range of $\approx 450$ Mpc. Assuming
the current best estimates for the cosmological parameters, this is equivalent
to a redshift $z\approx 0.1$. As noted by~\citet{1986Natur.323..310S},\ac{CBC}
systems can be used as cosmological distance markers, otherwise known as
``standard sirens'' (analogous to \ac{EM} standard candles). The nature of
these sources will allow us to estimate, amongst other parameters, their
luminosity distance without the need for calibration against the cosmic
distance ladder.

% BNS systems are GRBs
It is considered likely that the merger of \ac{BNS} and/or \ac{NSBH} systems,
in addition to emitting detectable \acp{GW}, is also the mechanism for
producing \acp{sGRB} \cite{1992ApJ...395L..83N}. In this scenario these events
produce tightly beamed \ac{EM} emission parallel to the orbital angular
momentum vector of the \ac{BNS} system. Additional evidence for the coincidence
of \acp{sGRB} with \ac{GW} events are the estimated astrophysical rates of both
phenomena. Observations of \acp{sGRB} give an inferred rate of \ac{sGRB}
producing \ac{CBC} mergers of $8\times 10^{-9}$--$1.1\times
10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$~\cite{2012MNRAS.425.2668C} which can be
compared to independently obtained estimates of \ac{BNS} merger rates of
$10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or
\ac{NSBH} merger rates of
$6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$
\cite{2010CQGra..27q3001A}. However, the recent work of
\cite{2015MNRAS.448..928K}, which studied known galactic \ac{BNS} systems in
more detail suggest a rate approximately five times lower. The estimates for
\ac{NSBH} systems come entirely from population synthesis models as no such
systems have yet been observed.
%The ranges of rates for \acp{BNS} is obtained from population synthesis models
%together with knowledge of the X known \ac{BNS} systems in our galaxy, whilst
%the range for \acp{NSBH} is entirely from population synthesis as not such
%systems have been observed.

@@

\MP{Move this paragraph to conclusions} For future planned space-based detectors such as eLISA there are several known galactic binary 
systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that their masses and 
orbital parameters allow them to be guaranteed sources for such detectors. These too could be used 
for assessing calibration for such detectors. However, generally for these verification binaries 
the inclination of the system is not known and distances are fairly uncertain, so an accurate 
calibration assessment may not be possible. Although there is at least one (J0651) currently known 
eclipsing white dwarf binary \cite{2041-8205-737-1-L23}, which does provide the system inclination 
\cite{2012ApJ...757L..21H, 2013ASPC..467...47K} and has an approximately 10\% distance uncertainty 
\cite{2041-8205-737-1-L23}, which make it an excellent candidate for an astrophysical calibrator at 
its frequency.

@@

% Standard siren basics
The idea of \ac{GW} standard sirens is directly analogous to the concept of
standard candles in \ac{EM} astronomy and was first proposed
in~\cite{1986Natur.323..310S}. Unlike the primary \ac{EM} standard candle
event, Type Ia Supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
not only a function of distance. It also depends upon the chirp mass (a
function of the component masses) and the binary orientation with respect to
the detector network. However, measurement of the phase evolution of such an
event allows accurate determination of the chirp mass~\footnote{In reality it
is the redshifted chirp mass that is measured.}. Timing information from the
different signal arrival times at each detector in the network allows
reasonably accurate sky position determination~\cite{2014ApJ...795..105S}.
Finally, amplitude variation between differently oriented interferometers
allows some (weaker) level of determination of the binary system orientation.
%A great help to this final stage for this project is the assumption that the
%source has an \ac{sGRB} counterpart.  It follows that we therefore have a well
%constrained prior belief on the inclination of the source orbit with respect
%to the line-of-sight based on the expected level of beaming in the \ac{sGRB}
%emission.

% The classic use of sirens
Standard sirens are clearly a very powerful tool for \ac{GW} cosmology since
they give us a direct measure of the absolute luminosity distance to sources.
This is distinct from Type Ia Supernovae standard candles that only provide
relative luminosity distance measures and require calibration via other methods
as part of the cosmological distance ladder. The use of \ac{GW} events with
\ac{sGRB} counterparts for use as a cosmological tool was investigated
in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
the 3rd generation \ac{GW} interferometer, the \ac{ET}. In this case, the
redshift information obtained from the \ac{sGRB} host galaxy, provides a
complementary measurement allowing each event to inform the distance-redshift
relationship and hence obtain cosmological parameter estimates. Other methods
for cosmological inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T,
2012PhRvD..86b3502T,Messenger:2011ux,2013arXiv1312.1862M}) have been proposed
for \ac{BNS} systems that do not rely on any redshift measurements. These
instead use either the distribution of measured \acp{SNR} and assumed \ac{NS}
mass and spatial distributions or exploit the features of the tidal and
post-merger hyper-massive \ac{NS} stages of the \ac{GW} waveform.

@@

% discuss joint-event rates
The predicted detection rates of \ac{BNS} events (irrespective of an \ac{sGRB}
counterpart) are derived from three main sources, population synthesis models,
extrapolations based on known galactic \ac{BNS} systems, and from observations
of \ac{sGRB} events. Various studies using the first two methods have been
compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{We note that this
rate is that given in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO}
and \ac{AdV} network consisting to two \ac{aLIGO} detectors operating at design
sensitivity. However, a higher rate of $0.4$--$400$ is given in
\cite{2010CQGra..27q3001A}, which may differ due to the use of a third
\ac{aLIGO} detector, but the discrepancy may also have contributions from
slightly different threshold definitions or design sensitivity curves used. We
also note the study mentioned in Sec.~\ref{sec:intro} that potentially lowers
the expected rate by a factor of five.}.  In~\cite{2012MNRAS.425.2668C} using
the rate of observed \acp{sGRB} and assuming a beaming angle of $15^{o}$ a
similar event rate of $1$--$180$ is obtained. The beaming angle is the primary
factor in determining the fraction of detected \ac{GW} \ac{BNS} signals with an
\ac{EM} \ac{sGRB} counterpart and is unfortunately highly uncertain.  Various
authors have provided estimates on the rate of joint detections to be in the
range $0.02$--$7$ per
year~\cite{2013PhRvL.111r1101C,Kelley:2012fl,2014arXiv1405.2254S,2015MNRAS.448.3026W}.

@@

% the main aim of the analysis
The main aim of this work is to assess how well the calibration scale factor
defined in Eq.~\ref{eq:scalefactor} can be estimated from a single observed
\ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have
little knowledge of the likely location and distance of such an event we have
performed simulations of multiple events to see the distribution in the accuracy of the calibration 
scale factor recovery.

@@

% mention of distance delta function prior
The primary reason why coincident observations with a \ac{sGRB} can be used as
a check of detector calibration is that the \ac{sGRB} allows us to constrain
the priors on various parameters that would normally be highly correlated with
the signal amplitude. The most important of these is that the \ac{sGRB} can
provide a very tight constraint on the source distance. In this analysis we therefore
assume that the source distance is known (i.e.\ has a $\delta$-function
prior) whereas in reality the error on this distance will be dependent
upon the uncertainties in both the redshift measurement and the Hubble
constant. For spectroscopic redshifts this uncertainty will be dominated by the
Hubble constant and will be of $O(1\%)$.

@@

% discuss the component mass and spin priors 
The prior on the system component masses is dependent on whether we are
considering the source being a \ac{BNS} system or a \ac{NSBH}. For the former
case the prior we use is a Gaussian distribution for both components with means
of $1.35\,\textrm{M}_{\odot}$ and standard deviations of
$0.13\,\textrm{M}_{\odot}$, which is roughly consistent with the double neutron star system
masses found in \cite{2012ApJ...757...55O} albeit with a wider distribution. In reality, for all 
the systems we use in our study, the chirp mass, which we see from Eq.~\ref{eq:signal}
plays a role in the overall signal amplitude, is very well constrained by the
system's phase evolution, so our prior could be expanded with minimal effect on
the results. In the case of \ac{NSBH} systems we use the same prior as above
for the neutron star, but for the black hole we use a prior based on the
canonical mass distribution used for the rate results in
\cite{2012PhRvD..85h2002A} with a mean of $5\,\textrm{M}_{\odot}$ and a
standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass range of
black holes could be quite different from this, but use this range as an
example for this type of system. For the \ac{NSBH} systems we also require a
prior on the black hole spin (for \ac{BNS} systems we have assumed that the
spins are small enough that they will be negligible). We use a uniform prior on
the normalised aligned spin magnitude between $-1$ and $1$.

@@

% Describe the ensemble of simulations used
Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
number (several hundred up to $O(1000)$) of simulations at each distance value.
When proposing injections for each simulation the source parameters were
randomly drawn from the prior distributions described in
Section~\ref{sec:priors}, with the sky position being drawn randomly from a
uniform distribution on the sky, the coalescence times held fixed within the
centre of the prior window, and with the calibration scaling factors for each
detector drawn from a Gaussian distribution with a mean of one and standard
deviation of $0.125$ (equivalent to a mean fractional offset of 10\%).  Note
that this distribution is broadly consistent with the scale of calibration
uncertainties that are expected via experimental methods but not the same as
the prior that we assume on $\mathcal{C}$ (see Eq.~\ref{eq:scale_prior}) when
performing our analysis.

% describe the detection cut 
In accepting a proposed injection as one to be analysed we introduced a
criteria that the signal be ``detectable'', based on that used in
\cite{2012PhRvD..85h2002A}. We therefore only analysed injections that had an
\ac{SNR} of $\geq 5.5$ in at least two detectors. Since we assume that these
signals would be coincident with a \ac{sGRB} we do not include the often used
further constraint that the total coherent network \ac{SNR} is greater than 12.
For injections close to the \ac{SNR} threshold, the application of the
threshold resulted in the preferential selection of better oriented systems.
Therefore, at larger distances the population of injected sources was not
uniformly distributed over the sky and instead favoured locations with better
antenna response.  Similarly, despite being already constrained by the prior to
small inclination angles, the accepted injections show preference for sources
very close to circularly polarised.

@@

% describe the main boxplot results
From these posterior distributions we have calculated the minimal 68\% credible
region for the calibration scale factors for each detector (if these were
Gaussian distributions this is equivalent to the region either side of the mean
bounded by the $1\sigma$ intervals). For all the signals at each distance
increment we have produced the distribution of the fractional half widths
(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the
true value. These are shown as boxplots in figures~\ref{fig:bnsresults} and
\ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems respectively. The
boxes show the extent from the lower to upper quartile of the values, whilst
the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The
black line within each box gives the median value and the star gives the mean
value. Also shown on each plot as the dashed magenta line is the percentage of
signals drawn from the prior distribution that fulfill the detection criterion.

@@

% discuss the BNS plots - general behaviour
In figure~\ref{fig:bnsresults} we see that on average the scale factor can be
recovered to equivalent precision in both \ac{aLIGO} detectors, as would be
expected, with uncertainties generally within the 10\% range for sources at
100\,Mpc. This is comparable to previous estimates of the calibration error in
the initial LIGO detectors. An interesting feature is that for distances
$\gtrsim 250$\,Mpc the upper extent of the uncertainty for H1 and L1 hits a
maximum at $\sim 25\%$, whilst the width of boxes narrows. This is due to our
``detectibility'' criteria, where for all distances we will only be seeing
those sources with a high enough \ac{SNR} that we would consider them
detectable. There will therefore be fewer sources with \ac{SNR} higher than
this criteria at large distances, giving us a narrower range, and we
automatically exclude those with \ac{SNR} that are too small thus truncating
our uncertainty distribution at the upper end.  However, this does show that on
average for sources that are detectable out to 450\,Mpc we would be able to
constrain the calibration scale factor uncertainty for the \ac{aLIGO} detectors
to $\sim 20\%$.

@@

% summary of main results
We have established how well we can assess the detector calibration for
\ac{aLIGO} and \ac{AdV} using astrophysical sources as standard sirens. To do
this we require that a \ac{sGRB}, for which it has been possible to measure the
distance, is observed in coincidence with a \ac{CBC} signal in the \ac{GW}
detectors. This enables us to assume a known distance and sky position for the
source, and also limit the inclination of the source, which in turn allows us
to test the consistency of the detector calibration. We do this by including an
unknown scale factor on the true signal and noise, which we estimate given the
data from the 3 interferometer advanced detector network. We find that for
detectable \ac{BNS} sources the uncertainty on the calibration scale factor
could on average be determined to $\lesssim 10\%$ of its true value for the
\ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the
proposed accuracy level of the hardware calibration of the detectors. For
sources at the standard single detector \ac{BNS} horizon distance of $\sim
450$\,Mpc the scale factor could on average be determined to within $\lesssim
20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were
found for \ac{NSBH} sources, although sources could be observed out to higher
distances.

@@

% comparative calibration
%We imagine that a test for consistency of calibration between detectors could be
%performed with an ensemble of \ac{CBC} sources without relying on \ac{EM}
%counterparts. Such a method would have to marginalise over highly
%uncertain population models and may give quite weak bounds on the overall
%calibration uncertainty.~\cm{I'm not sure what you mean here? I want to add a
%paragraph on relative calibration after I understand what you mean.}

% comparative calibration
A further application of astrophysical calibration is the relative calibration
of \ac{GW} detectors.  In this case we would be interested in obtaining
posterior distributions on the ratios of calibration scale factors between
detectors. For \ac{CBC} events, in this scenario there would be no explicit
requirement for a redshift measurement and therefore a cosmologically inferred
distance. There would however be a requirement for an electromagnetically
inferred sky position of $O(\rm{degrees})$ accuracy. A more suitable source for
relative calibration estimation would be that of continuous \ac{GW} emission.
In that case, sky position and orientation information would be obtained to a
high level of accuracy from single detector analyses alone. A primary use for
relative calibration information would be to account for potential biases in
the sky position estimates for \ac{CBC} and burst sources. In this case,
amplitude information is a secondary, but important factor, in breaking
degeneracies on sky position estimates.

% discuss PSD uncertainty
We emphasise that our work has assumed that the \ac{PSD} used in the likelihood
function is an accurate representation of the \ac{PSD} at the time of the
observed \ac{CBC} signal. In reality the \ac{PSD} may be estimated from a
period close to, but not overlapping, the signal time and therefore may be
slightly different \cite{2013PhRvD..88h4044L}. Our result would therefore be
highly correlated with any uncertainty in the \ac{PSD} estimate, but would
still offer an upper limit on the calibration uncertainty. 
%However, future
%plans for \ac{CBC} parameter estimation include estimating the \ac{PSD} with
%the signal using BayesLine \cite{2015PhRvD..91h4034L}~\cm{I'm nit sure that
%this last point really helps us.}.

% frequency dependence
A further extension of this work would be for the calibration scale factor to
be estimated as a function of frequency. This may require the frequency series
to be divided into sections and the calibration scale factor estimated for
each. However, since the scale factor uncertainty is related to the \ac{SNR}
then the \ac{SNR} contribution in each frequency section will be of relevance.
In this case Bayesian model selection may be applicable for selecting the
optimal number of frequency sections with which to estimate the scale factor.
The BayesLine algorithm \cite{2015PhRvD..91h4034L}, or the method described in
\cite{2013PhRvD..88h4044L}, may provide a natural way to perform such an
analysis. Indeed these methods, and those of \cite{Vitale:2012} are already
being used for marginalising over calibration uncertainties, but do not yet
attempt to estimate it.

@@

We would like to acknowledge useful discussions within the LIGO-Virgo
Collaboration and specifically with Salvatore Vitale. LW was part funded for
this work through a Royal Astronomical Society Undergraduate Research Bursary.
MP and CM are funded by the Science and Technology Research Council under grant
number ST/L000946/1. CM is also supported by a Glasgow University Lord Kelvin
Adam Smith Fellowship. This document has LIGO DCC number \dcc.
